Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-12 doc/report/msc-report.tex
Introduction: background.

Any organization will eventually wish to evaluate and analyze various
characteristics of the service and operations they manage. These services can
be anything from web sites to theme parks, and the evaluation might include
characteristics such as the number of visitors per day, and the average time a
theme park guest spends queueing for a particular ride.

Inevitably, these services and operations will generate a log for later
analysis. As the services grow in size, so will the data sets they generate.

An important property of these data sets is that they generally have a regular
structure and are represented as a number of independent records. The records
can be broken up without losing any of their meaning and are as such very well
suited for parallel analysis. The data sets are not fit to store using
traditional database techniques, if only for the fact that they are too large
to fit in a relational database.

Each log record typically contains quite a bit of information. For example, a
log of visits to a web site might contain a timestamp, information about what
country the visiting user is from, and a string of text that identifies the
web browser that was used. From these base facts one can collect information
about what countries generate the most traffic to the site, and which web
browser is most popular in a certain country, for example.

These bits of information can be thought of as points along an axis; Sweden,
Norway and Denmark are all points along the country axis, for instance. This
is an important concept and it will be used throughout the text.

Typically, the size of a data set is in the range of several gigabytes of data
per day. A single computer is not likely to be able to process that amount of
data before a new batch arrives, and for very large data sets the single
computer would simply not be able to fit the intermediate data set in its
memory at all. Some data sets are many times larger than that and do not even
fit inside a single computers storage devices.

>>>>

Any organization will eventually wish to analyze various characteristics of
its service and operations. These services can be anything from web sites to
theme parks. The analysis might include characteristics such as the number of
visitors per day, or the average time a theme park guest spends queueing for a
particular ride.

Inevitably, the services and operations will generate a log for later
analysis. As the services grow in size, so will the data sets they generate.

An important property of these data sets is that they generally have a regular
structure and are broken down into records that are independent of each other.
The records can be analyzed one at a time without losing any of their meaning
and are as such very well suited for parallel analysis.

Each log record typically contains a number of fields. For example, a log of
visits to a web site might contain a timestamp, information about what country
the visitor is from and what web browser she used. The values of the fields
can be thought of as points along an axis; Sweden, Norway and Denmark are all
points along the country axis, for instance. This is an important concept and
it will be touched upon repeatedly throughout the text.

After having collected these base facts, one can draw conclusions about what
countries generate the most traffic to the site, and which web browser is most
popular in a certain country, and so on.

Typically, the size of a data set is in the range of several gigabytes of data
per day. Some data sets are many times larger.

A single computer is not likely to be able to process that amount of data
before a new batch arrives, and for very large data sets a single computer
would simply not be able to fit the intermediate data set in its working
memory.
